---
title: "Make 'Source data file'"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    rmarkdown::github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "md_files/source_data-",
  out.width = "100%"
)
reRun <- FALSE
```

```{r}
library(tidyverse)
library(TPP2D)
library(readxl)
library(openxlsx)
```

# Main figures

## Figure 1:               
- schematic illustrations only              
        
        
## Figure 2 - Benchmark        

```{r}
sim_tp_df <- readRDS("../pre_run_data/sim_tp_df.rds")
sim_params_df <- readRDS("../pre_run_data/sim_params_df.rds")
sim_fstat_df <- computeFStatFromParams(sim_params_df)
sim_s1_null_df <- readRDS("../pre_run_data/sim_s1_null_df.rds")
sim_s2_null_df <- readRDS("../pre_run_data/sim_s2_null_df.rds")
sim_s3_null_df <- readRDS("../pre_run_data/sim_s3_null_df.rds")
thres_found_recurrently <- readRDS("../pre_run_data/thres_found_recurrently.rds")
```

Define function to compute the precision and recall based on the simulated dataset:

```{r}
computePrecisionRecall <- function(fdr_df,
                                   alpha_range = seq(0.01, 1, by = 0.01),
                                   ntp = 50){
  out_pr_df <- bind_rows(tibble(precision = 1, recall = 0, alpha = 0),
                         bind_rows(lapply(alpha_range, function(alpha_i){
    hits_df <- TPP2D::findHits(fdr_df, alpha = alpha_i)
    
    r_df <- tibble(
        precision = length(which(!grepl("protein", hits_df$clustername)))/
            nrow(hits_df),
        recall = length(which(!grepl("protein", hits_df$clustername)))/
            ntp,alpha = alpha_i)
  })))
  
  return(out_pr_df)
}

```

Get FDR estimates for different numbers of bootstraps
```{r}
sim_fdr_df_s1_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s1_null_df,
                                     squeezeDenominator = FALSE)
sim_fdr_df_s2_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s2_null_df,
                                     squeezeDenominator = FALSE)
sim_fdr_df_s3_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s3_null_df,
                                     squeezeDenominator = FALSE)

sim_fdr_df_s1_b100_byMsExp_mod <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s1_null_df,
                                     squeezeDenominator = TRUE)
sim_fdr_df_s2_b100_byMsExp_mod <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s2_null_df,
                                     squeezeDenominator = TRUE)
sim_fdr_df_s3_b100_byMsExp_mod <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s3_null_df,
                                     squeezeDenominator = TRUE)
```

Compute precision and recall for each of the FDR estimates based on the different number of bootstraps
```{r}
pr_b100_1_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s1_b100_byMsExp, ntp = 80)
pr_b100_2_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s2_b100_byMsExp, ntp = 80)
pr_b100_3_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s3_b100_byMsExp, ntp = 80)

pr_b100_1_byMsExp_mod <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s1_b100_byMsExp_mod, ntp = 80)
pr_b100_2_byMsExp_mod <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s2_b100_byMsExp_mod, ntp = 80)
pr_b100_3_byMsExp_mod <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s3_b100_byMsExp_mod, ntp = 80)

all_alpha_fdr_df <- bind_rows(
  pr_b100_1_byMsExp %>% 
     mutate(variant = "DLPTP standard, seed 1"), 
   pr_b100_2_byMsExp %>% 
     mutate(variant = "DLPTP standard, seed 2"), 
   pr_b100_3_byMsExp %>% 
     mutate(variant = "DLPTP standard, seed 3"), 
  pr_b100_1_byMsExp_mod %>% 
     mutate(variant = "DLPTP moderated, seed 1"),
   pr_b100_2_byMsExp_mod %>% 
     mutate(variant = "DLPTP moderated, seed 2"),
   pr_b100_3_byMsExp_mod %>% 
     mutate(variant = "DLPTP moderated, seed 3")) %>% 
  group_by(variant, alpha) %>% 
  summarize(mean_fdr = mean(1-precision),
            mean_tpr = mean(recall)) %>% 
  ungroup


ggplot(all_alpha_fdr_df, aes(alpha, mean_fdr)) +
  geom_point(aes(color = variant)) +
  geom_path(aes(color = variant)) +
  geom_abline(slope = 1, 
              color = "darkgray",
              linetype = "dashed") +
  coord_cartesian(xlim = c(0, 0.2),
                  ylim = c(0, 0.2)) +
  scale_color_manual(
    "variant",
    values = c("deepskyblue4",
               "deepskyblue3",
               "deepskyblue1",
               "darkorchid4",
               "darkorchid3",
               "darkorchid1")) +
  labs(x = "nominal FDR",
       y = "observed FDR") +
  theme_classic()

```


```{r}
b100_pr_df <- bind_rows(
  pr_b100_1_byMsExp %>% 
    mutate(variant = "DLPTP standard"), 
  pr_b100_2_byMsExp %>% 
    mutate(variant = "DLPTP standard"),
  pr_b100_3_byMsExp %>% 
    mutate(variant = "DLPTP standard"), 
  pr_b100_1_byMsExp_mod %>% 
    mutate(variant = "DLPTP moderated"),
   pr_b100_2_byMsExp_mod %>% 
    mutate(variant = "DLPTP moderated"),
  pr_b100_3_byMsExp_mod %>% 
    mutate(variant = "DLPTP moderated"),
  tibble(
    precision = length(thres_found_recurrently[
      !grepl("protein",thres_found_recurrently)])/
      length(thres_found_recurrently),
    recall = length(thres_found_recurrently[
      !grepl("protein", thres_found_recurrently)])/80,
    alpha = 0,
    variant = "Threshold-based approach")) %>% 
  group_by(variant, alpha) %>% 
  summarize(mean_fdr = mean(1-precision),
            mean_tpr = mean(recall)) %>% 
  ungroup 
  

ggplot(b100_pr_df, aes(x = mean_fdr, y = mean_tpr, 
                       color = as.factor(variant))) + 
  geom_path(size = 0.5) + 
  geom_point(shape = 8, size = 3,
             data = filter(filter(b100_pr_df, alpha == "0.1"))) +
    geom_point(shape = 3, size = 3,
             data = filter(filter(b100_pr_df, alpha == "0.05"))) +
    geom_point(shape = 4, size = 3,
             data = filter(filter(b100_pr_df, alpha == "0.01"))) +
  geom_point(shape = 6, size = 3,
             data = filter(filter(b100_pr_df, 
                                  variant == "Threshold-based approach"))) +
  scale_color_manual("variant", 
                     values = c("deepskyblue4", "darkorchid4", "darkorange1")) +
  labs(x = "FDR", y = "TPR") +
  geom_vline(xintercept = c(0.01, 0.05, 0.1),
             linetype = "dashed", alpha = 0.65,
             size = 0.25) +
  coord_fixed() +
  theme_classic() +
  theme(legend.position = c(0.75, 0.2))

```

## Figure 3 - Panobinostat & JQ1

### Panobinostat
```{r}
pano_cell_raw <- read_xlsx(
    file.path("..", "Panobinostat_HepG2_cell_2DTPP_Becher_et_al_2016",
              "41589_2016_BFnchembio2185_MOESM254_ESM.xlsx"), 
                           sheet = 1, skip = 1) %>% 
  dplyr::select(representative,
                clustername,
                experiment = ms_experiment,
                qupm,
                qusm,
                temperature,
                matches("sumionarea"),
                -matches("total"),
                matches("rel_fc_protein"),
                -matches("transformed"),
                -matches("orig"),
                -matches("log2rel"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
pano_cell_fil <- resolveAmbiguousProteinNames(pano_cell_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
pano_cell_df <- recomputeSignalFromRatios(pano_cell_fil)
```

```{r}
pano_params_df <- readRDS("../pre_run_data/pano_params_df.rds")
```

Compute *F* statistics
```{r}
pano_fstat_df <- computeFStatFromParams(pano_params_df)
```

Get $B$,prerun datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:

```{r}
pano_null_df <- readRDS("../pre_run_data/pano_null_df.rds")
```

Compute FDR and find hits:

```{r}
pano_fdr_df <- getFDR(df_out = pano_fstat_df,
                     df_null = pano_null_df,
                     squeezeDenominator = TRUE)
  
pano_hits_df <- findHits(pano_fdr_df, alpha = 0.1)
```

Make output table
```{r}
pano_out_tab <- pano_fdr_df %>% filter(dataset == "true") %>% 
    dplyr::select(protein_id = representative, 
                  gene_name = clustername, nObsRound, 
                  rssH0,  rssH1, F_statistic) %>% 
    mutate(stabilized_hit_at_10percent_FDR = 
               case_when(
                   protein_id %in% 
                       filter(pano_hits_df, slopeH1 > 0)$representative ~ 
                       TRUE, TRUE ~ FALSE),
           destabilized_hit_at_10percent_FDR = 
               case_when(protein_id %in% 
                             filter(pano_hits_df, slopeH1 < 0)$representative ~
                             TRUE, TRUE ~ FALSE)) 
```

Make sure we can still make the plot
```{r}
ggplot(pano_out_tab, aes(log2(rssH0-rssH1), asinh(F_statistic))) +
    geom_point(color = "gray", alpha = 0.5) +
    geom_point(color = "steelblue", alpha = 0.5, 
               data = filter(pano_out_tab, stabilized_hit_at_10percent_FDR)) +
    geom_point(color = "orange", alpha = 0.5, 
               data = filter(pano_out_tab, destabilized_hit_at_10percent_FDR)) +
    coord_cartesian(xlim = c(-12.5, 7.5))
```

```{r}
pano_thres_df <- read_xlsx("../Panobinostat_HepG2_cell_2DTPP_Becher_et_al_2016/41589_2016_BFnchembio2185_MOESM254_ESM.xlsx", 
                          sheet = 1, skip = 1) %>% 
  filter(qupm > 1)

all_pano_hits <- union(
  (pano_thres_df %>% 
     filter(protein_stabilized_neighb_temp_good_curves_count > 1 | 
              protein_destabilized_neighb_temp_good_curves_count > 1) %>% 
     filter(!duplicated(clustername)))$clustername,
  pano_hits_df$clustername)

pano_venn_df <- tibble(
  gene_name = all_pano_hits) %>% mutate(
    DLPTP_stabilized_hits = as.numeric(
      gene_name %in% filter(pano_hits_df, slopeH1 > 0)$clustername),
    DLPTP_destabilized_hits = as.numeric(
      gene_name %in% filter(pano_hits_df, slopeH1 < 0)$clustername),
    Threshold_based_stabilized_hits = as.numeric(
      gene_name %in% 
        filter(pano_thres_df, protein_stabilized_neighb_temp_good_curves_count > 1)$clustername),
    Threshold_based_destabilized_hits = as.numeric(
      gene_name %in% 
        filter(pano_thres_df, protein_destabilized_neighb_temp_good_curves_count > 1)$clustername))
```


### JQ1

```{r}
jq1_lys_raw <- read_xlsx(file.path(
  "../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/", 
  "Savitski_et_al_Figure_3/Supplementary Dataset 2_2D-TPP.xlsx"), 
  sheet = 3, skip = 1) %>% 
  dplyr::select(representative = `Accession No.`,
                clustername = `protein name`,
                qupm = QUPM,
                qusm = QUSM,
                temperature,
                matches("sumionarea"),
                -matches("total"),
                matches("rel_fc_protein"),
                -matches("transformed"),
                -matches("orig"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
jq1_lys_fil <- resolveAmbiguousProteinNames(jq1_lys_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
jq1_lys_df <- recomputeSignalFromRatios(jq1_lys_fil)
```

Compute null and alternative model fits and extract parameters
```{r eval=reRun}
jq1_params_df <- getModelParamsDf(jq1_lys_df, maxit = 500)
saveRDS(jq1_params_df, file = "../pre_run_data/jq1_params_df.rds")
```

```{r}
jq1_params_df <- readRDS("../pre_run_data/jq1_params_df.rds")
```

Compute *F* statistics
```{r}
jq1_fstat_df <- computeFStatFromParams(jq1_params_df)
```

Get $B$, prerun datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:

```{r}
jq1_null_df <- readRDS("../pre_run_data/jq1_null_df.rds")
```

Compute FDR and find hits:

```{r}
jq1_fdr_df <- getFDR(df_out = jq1_fstat_df,
                     df_null = jq1_null_df,
                     squeezeDenominator = TRUE)
  
jq1_hits_df <- findHits(jq1_fdr_df, alpha = 0.1)
```

Make output table

```{r}
jq1_out_tab <-  
    jq1_fdr_df %>% 
        filter(dataset == "true") %>% 
        dplyr::select(protein_id = representative, 
                      gene_name = clustername, nObsRound, 
                      rssH0,  rssH1, F_statistic) %>% 
    mutate(stabilized_hit_at_10percent_FDR = 
               case_when(protein_id %in% 
                             filter(jq1_hits_df, slopeH1 > 0)$representative ~
                             TRUE, TRUE ~ FALSE),
           destabilized_hit_at_10percent_FDR = 
               case_when(protein_id %in% 
                             filter(jq1_hits_df, slopeH1 < 0)$representative ~
                             TRUE, TRUE ~ FALSE))
```

Make sure we can still make the plot
```{r}
ggplot(jq1_out_tab, aes(log2(rssH0-rssH1), asinh(F_statistic))) +
    geom_point(color = "gray", alpha = 0.5) +
    geom_point(color = "steelblue", alpha = 0.5, 
               data = filter(jq1_out_tab, stabilized_hit_at_10percent_FDR)) +
    geom_point(color = "orange", alpha = 0.5, 
               data = filter(jq1_out_tab, destabilized_hit_at_10percent_FDR)) +
    coord_cartesian(xlim = c(-12.5, 7.5))
```

```{r}
jq1_thres_df <- read_xlsx("../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3/Supplementary Dataset 2_2D-TPP.xlsx", sheet = 3, skip = 1) %>% 
  filter(QUPM > 1)

all_jq1_hits <- union(
  (jq1_thres_df %>% 
     filter(protein_stabilized_neighb_temp_good_curves_count > 1 | 
              protein_destabilized_neighb_temp_good_curves_count > 1) %>% 
     filter(!duplicated(`protein name`)))$`protein name`,
  jq1_hits_df$clustername)

jq1_venn_df <- tibble(
  gene_name = all_jq1_hits) %>% mutate(
    DLPTP_stabilized_hits = as.numeric(
      gene_name %in% filter(jq1_hits_df, slopeH1 > 0)$clustername),
    DLPTP_destabilized_hits = as.numeric(
      gene_name %in% filter(jq1_hits_df, slopeH1 < 0)$clustername),
    Threshold_based_stabilized_hits = as.numeric(
      gene_name %in% 
        filter(jq1_thres_df, protein_stabilized_neighb_temp_good_curves_count > 1)$`protein name`),
    Threshold_based_destabilized_hits = as.numeric(
      gene_name %in% 
        filter(jq1_thres_df, protein_destabilized_neighb_temp_good_curves_count > 1)$`protein name`),
    )
```

## Figure 4 - PCI-34051 and BRD-3811


###  PCI-34051
```{r}
pci_raw <- read_xlsx("../PCI-34051_HL60_cell_2DTPP/Supplementary_Data_2.xlsx", sheet = "PCI34051") %>% 
  dplyr::select(representative,
                clustername,
                qupm,
                qusm,
                temperature,
                matches("sumionarea"),
                matches("rel_fc_protein"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
pci_fil <- resolveAmbiguousProteinNames(pci_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
pci_df <- recomputeSignalFromRatios(pci_fil)
```

```{r}
pci_params_df <- readRDS("../pre_run_data/pci_params_df.rds")
```

Compute *F* statistics
```{r}
pci_fstat_df <- computeFStatFromParams(pci_params_df)
```

Get $B$, prerun datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:

```{r}
pci_null_df <- readRDS("../pre_run_data/pci_null_df.rds")
```

```{r}
## manually identified carry-over cases
carry_over_cases <- 
  c("ALDH1B1", "BTK", "CAMK2G", "CSK", "CSNK2A1", "CSNK2A2", "GAK",
    "CSNK2B", "GSK3A", "LYN", "MAP4K1", "MAPK1", "MAPK9", "NEK9", 
    "NQO1", "PDXK", "PRKAA1", "PRKAG1", "RPS6KA1", "ULK3",
    "CDK2", "CDK5")

pci_fstat_df <- computeFStatFromParams(pci_params_df) %>% 
  filter(!clustername %in% carry_over_cases)
pci_fdr_df <- getFDR(df_out = pci_fstat_df,
                     df_null = pci_null_df %>% 
                       filter(!clustername %in% carry_over_cases))

pci_hits_df <- findHits(pci_fdr_df, alpha = 0.1)


pci_out_tab <- pci_fdr_df %>% 
    filter(dataset == "true") %>% 
    dplyr::select(protein_id = representative, 
                  gene_name = clustername, 
                  nObsRound, rssH0,  rssH1, F_statistic) %>% 
  mutate(affected_by_carry_over = case_when(
    gene_name %in% carry_over_cases ~ TRUE, TRUE ~ FALSE),
    stabilized_hit_at_10percent_FDR = case_when(
      protein_id %in% 
        filter(pci_hits_df, slopeH1 > 0)$representative ~ TRUE, 
      TRUE ~ FALSE),
    destabilized_hit_at_10percent_FDR = case_when(
      protein_id %in% 
        filter(pci_hits_df, slopeH1 < 0)$representative ~ TRUE, 
      TRUE ~ FALSE))
```

Make sure we can still make the plot
```{r}
ggplot(pci_out_tab, aes(log2(rssH0-rssH1), asinh(F_statistic))) +
    geom_point(color = "gray", alpha = 0.5) +
    geom_point(color = "steelblue", alpha = 0.5, 
               data = filter(pci_out_tab, stabilized_hit_at_10percent_FDR)) +
    geom_point(color = "orange", alpha = 0.5, 
               data = filter(pci_out_tab, destabilized_hit_at_10percent_FDR)) +
    coord_cartesian(xlim = c(-12.5, 7.5))
```

### BRD-3811

```{r}
brd_raw <- read_xlsx("../BRD-3811_HL60_cell_2DTPP/Supplementary_Data_2.xlsx", sheet = "BRD3811") %>% 
  dplyr::select(representative,
                clustername,
                qupm,
                qusm,
                temperature,
                matches("sumionarea"),
                matches("rel_fc_protein"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
brd_fil <- resolveAmbiguousProteinNames(brd_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
brd_df <- recomputeSignalFromRatios(brd_fil)
```

Compute null and alternative model fits and extract parameters

```{r}
brd_params_df <- readRDS("../pre_run_data/brd_params_df.rds")
```

Compute *F* statistics
```{r}
brd_fstat_df <- computeFStatFromParams(brd_params_df)
```

Get $B$, prerun datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:

```{r}
brd_null_df <- readRDS("../pre_run_data/brd_null_df.rds")
```

```{r}
## manually identified carry-over cases
carry_over_cases_brd3811 <- c("NQO1", "CSK")

brd_fdr_df <- getFDR(df_out = brd_fstat_df %>% 
                         filter(!clustername %in% carry_over_cases_brd3811),
                     df_null = brd_null_df %>% 
                         filter(!clustername %in% carry_over_cases_brd3811),
                     squeezeDenominator = TRUE)
  
brd_hits_df <- findHits(brd_fdr_df, alpha = 0.1)
```


```{r}
brd_out_tab <- brd_fdr_df %>% filter(dataset == "true") %>%
    dplyr::select(protein_id = representative, 
                  gene_name = clustername, nObsRound,
                  rssH0,  rssH1, F_statistic) %>% 
  mutate(affected_by_carry_over = case_when(
    gene_name %in%
      carry_over_cases_brd3811 ~ TRUE, TRUE ~ FALSE),
    stabilized_hit_at_10percent_FDR = case_when(
      protein_id %in% 
        filter(brd_hits_df, slopeH1 > 0)$representative ~ TRUE, 
      TRUE ~ FALSE),
    destabilized_hit_at_10percent_FDR = case_when(
      protein_id %in% 
        filter(brd_hits_df, slopeH1 < 0)$representative ~ TRUE, 
      TRUE ~ FALSE))
```

Make sure we can still make the plot
```{r}
ggplot(brd_out_tab, aes(log2(rssH0-rssH1), asinh(F_statistic))) +
    geom_point(color = "gray", alpha = 0.5) +
    geom_point(color = "steelblue", alpha = 0.5, 
               data = filter(brd_out_tab, stabilized_hit_at_10percent_FDR)) +
    geom_point(color = "orange", alpha = 0.5, 
               data = filter(brd_out_tab, destabilized_hit_at_10percent_FDR)) +
    coord_cartesian(xlim = c(-12.5, 7.5))
```


### In vitro assay

```{r}
lap3_tab <- read_xlsx("LAP3_assay_simplified.xlsx") %>% 
  dplyr::select(`Time [s]`, matches("DMSO"), matches("100uM")) %>% 
  gather(key, value, -`Time [s]`) %>% 
  mutate(condition = factor(case_when(grepl("DMSO", key) ~ "control",
                                      grepl("BRD", key) ~ "100 uM BRD3811",
                                      TRUE ~ "100 uM PCI-34051"),
                            levels = c("control", "100 uM PCI-34051", "100 uM BRD3811"))) %>%
  group_by(`Time [s]`) %>% 
  mutate(replicate = factor(case_when(!duplicated(condition) ~ 1,
                                      TRUE ~ 2))) %>% 
  ungroup 

ggplot(lap3_tab, aes(`Time [s]`, value, color = condition)) +
  geom_point(aes(shape = replicate), size = 0.5) +
  stat_summary(fun = base::mean, geom = "line", size = 0.5) +
  ylab("Intensity") +
  scale_color_manual(values = c("black", "darkblue", "firebrick"))
```

## Figure 5

### GTP

Annotate GTP and ATP binding proteins

```{r}
all_atp_binder <- AnnotationDbi::select(
  org.Hs.eg.db::org.Hs.eg.db, 
  keys = "GO:0005524", 
  columns = c("SYMBOL", "IPI"), 
  keytype = "GOALL")

all_gtp_binder <- AnnotationDbi::select(
  org.Hs.eg.db::org.Hs.eg.db, 
  keys = "GO:0005525", 
  columns = c("SYMBOL", "IPI"), 
  keytype = "GOALL")
```

```{r}
gtp_raw <- read_xlsx("../GTP_Jurkat_gelfil_lysate_2DTPP/Supplementary_Data_3.xlsx", sheet = "GTP") %>% 
  dplyr::select(representative,
                clustername,
                qupm,
                qusm,
                experiment,
                temperature,
                matches("sumionarea"),
                matches("rel_fc_protein"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
gtp_fil <- resolveAmbiguousProteinNames(gtp_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
gtp_df <- recomputeSignalFromRatios(gtp_fil)
```


```{r eval=!reRun, echo=FALSE}
gtp_params_df <- readRDS("../pre_run_data/gtp_params_df.rds")
```

Compute *F* statistics
```{r}
gtp_fstat_df <- computeFStatFromParams(gtp_params_df)
```

Get $B$, prerun datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:

```{r eval=!reRun, echo=FALSE}
gtp_null_df <- readRDS("../pre_run_data/gtp_null_df.rds")
```

Compute FDR and find hits:

```{r}
gtp_fdr_df <- getFDR(df_out = gtp_fstat_df,
                     df_null = gtp_null_df,
                     squeezeDenominator = TRUE)
  
gtp_hits_10per <- findHits(gtp_fdr_df, 0.1) 
gtp_hits_1per <- findHits(gtp_fdr_df, 0.01)


gtp_out_tab <- gtp_fdr_df %>% filter(dataset == "true") %>% 
    dplyr::select(protein_id = representative, 
                  gene_name = clustername, nObsRound,
                  rssH0, rssH1, F_statistic)%>% 
  mutate(stabilized_hit_at_1percent_FDR = 
           case_when(protein_id %in% 
                       filter(gtp_hits_1per, slopeH1 > 0)$representative ~ TRUE,
                     TRUE ~ FALSE),
         destabilized_hit_at_1percent_FDR = 
           case_when(gene_name %in% 
                       filter(gtp_hits_1per, slopeH1 < 0)$clustername ~ TRUE,
                     TRUE ~ FALSE),
         stabilized_hit_at_10percent_FDR = 
           case_when(gene_name %in% 
                       filter(gtp_hits_10per, slopeH1 > 0)$clustername ~ TRUE,
                     TRUE ~ FALSE),
         destabilized_hit_at_10percent_FDR = 
           case_when(gene_name %in% 
                       filter(gtp_hits_10per, slopeH1 < 0)$clustername ~ TRUE,
                     TRUE ~ FALSE),
         GO_annotated_ATP_binder = case_when(
           gene_name %in% all_atp_binder$SYMBOL ~ TRUE, 
           TRUE ~ FALSE),
         GO_annotated_GTP_binder = case_when(
           gene_name %in% all_gtp_binder$SYMBOL ~ TRUE, 
           TRUE ~ FALSE))
```

Make sure we can still make the plot
```{r}
ggplot(gtp_out_tab, aes(log2(rssH0-rssH1), asinh(F_statistic))) +
    geom_point(color = "gray", alpha = 0.5) +
    geom_point(color = "black", alpha = 0.5, 
               data = filter(gtp_out_tab, stabilized_hit_at_10percent_FDR |
                               destabilized_hit_at_10percent_FDR)) +
    geom_point(color = "darkgreen", alpha = 0.5, 
               data = filter(gtp_out_tab, GO_annotated_ATP_binder &
                               (stabilized_hit_at_10percent_FDR |
                               destabilized_hit_at_10percent_FDR))) + 
    geom_point(color = "violet", alpha = 0.5, 
               data = filter(gtp_out_tab, GO_annotated_GTP_binder &
                               (stabilized_hit_at_10percent_FDR |
                               destabilized_hit_at_10percent_FDR))) +
    coord_cartesian(xlim = c(-12.5, 7.5)) +
    facet_wrap(~nObsRound)
```

```{r}
gtp_threshold_df <- read_xlsx("../GTP_Jurkat_gelfil_lysate_2DTPP/Supplementary_Data_3.xlsx", sheet = "GTP") %>% 
  dplyr::select(clustername, stabilized_hits_found_by_threshold) %>% 
  filter(stabilized_hits_found_by_threshold)

gtp_hits_10per <- findHits(gtp_fdr_df, 0.10) %>% 
  filter(detected_effectH1 == "stability", slopeH1 > 0)
gtp_hits_1per <- findHits(gtp_fdr_df, 0.01) %>% 
  filter(detected_effectH1 == "stability", slopeH1 > 0)


set_intersect_df <- data.frame(
  gene_name = filter(gtp_fdr_df, !duplicated(clustername))$clustername) %>% 
  mutate(`ATP binders` = as.numeric(gene_name %in% all_atp_binder$SYMBOL),
         `GTP binders` = as.numeric(gene_name %in% all_gtp_binder$SYMBOL),
         `DLPTP hits at 10% FDR` = 
           as.numeric(gene_name %in% gtp_hits_10per$clustername),
         `DLPTP hits at 1% FDR` = 
           as.numeric(gene_name %in% gtp_hits_1per$clustername),
         `Hits found by threshold-based approach` = 
           as.numeric(gene_name %in% gtp_threshold_df$clustername))

UpSetR::upset(set_intersect_df, nsets = 5)
```

# Supplementary figures

## Figure S1

```{r}
brd4_jq1_df <- filter(jq1_lys_df, clustername == "BRD4") %>% 
  dplyr::select(gene_name = clustername, qupm, qusm, 
                temperature, conc, log2_value, rel_value)
```


## Figure S2

### a == Table for Figure 3a

### b == Table for Figure 3b

### Panobinostat Venn Diagramms

### JQ1 Venn Diagramms

## Figure S3

```{r}
examples_pano_df <- filter(pano_cell_df, clustername %in%
                             c("FADS1", "HDAC1", "ZNF384",
                               "DHRS1", "HDAC6")) %>% 
  dplyr::select(gene_name = clustername, qupm, qusm, 
                temperature, conc, log2_value, rel_value)
```

## Figure S4

### a == Table for Figure 4a

### b == Table for Figure 4b

### c

```{r}
csk_pci_df <- filter(pci_df, clustername == "CSK") %>% 
  dplyr::select(gene_name = clustername, qupm, qusm, 
                temperature, conc, log2_value, rel_value)
```


# Make final list
```{r}
out_list <- list(
  "Figure2a" = all_alpha_fdr_df %>% as.data.frame(),
  "Figure2b" = b100_pr_df %>% as.data.frame(),
  "Figure3a_FigureS2a" = pano_out_tab %>% as.data.frame(),
  "Figure3b_FigureS2b" = jq1_out_tab %>% as.data.frame(),
  "Figure4b_FigureS4a" = pci_out_tab %>% as.data.frame(),
  "Figure4d_FigureS4b" = brd_out_tab %>% as.data.frame(),
  "Figure4e" = lap3_tab %>% as.data.frame(),
  "Figure5a" = gtp_out_tab %>% as.data.frame(),
  "Figure5b" = set_intersect_df %>% as.data.frame(),
  "FigureS1" = brd4_jq1_df %>% as.data.frame(),
  # "FigureS2a" = pano_out_tab,
  # "FigureS2b" = jq1_out_tab,
  "FigureS2c" = pano_venn_df %>% as.data.frame(),
  "FigureS2d" = jq1_venn_df %>% as.data.frame(),
  "FigureS3a-e" = examples_pano_df %>% as.data.frame(),
  "FigureS4c" = csk_pci_df %>% as.data.frame()
)
```

Write excel table
```{r}
library(openxlsx)
write.xlsx(out_list, col.names = TRUE, row.names = FALSE, asTable = TRUE, 
           file = "Kurzawa_et_al_source_data.xlsx")

```

