---
title: "Analysis of GTP dataset"
author:
- name: Nils Kurzawa
  affiliation: 
  - European Molecular Biology Laboratory (EMBL), Genome Biology Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
package: TPP2D
output:
  BiocStyle::pdf_document:
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8} 
bibliography: ../bibliography.bib
csl: ../cell.csl
header-includes: 
- \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reRun <- FALSE
```

# Step-by-step walk through the anlysis

```{r eval = FALSE}
# This script uses the development version of TPP2D
if(require("BiocManager"))
  install.packages("BiocManager")
BiocManager::install("nkurzaw/TPP2D")
```

Load required libraries

```{r}
library(TPP2D)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(UpSetR)
```

Define plot style
```{r}
theme_paper <- theme_bw(base_size = 6) +
  theme(legend.background = element_blank(), 
        legend.key = element_blank(), 
        panel.background = element_blank(), 
        panel.grid.major = element_line(colour = "grey92", size = 0.25),
        panel.grid.minor = element_line(colour = "grey92", size = 0.15),
        panel.border = element_blank(), 
        strip.background = element_blank(), 
        plot.background = element_blank(), 
        complete = TRUE,
        axis.line = element_line(color = "black", size = 0.25),
        text = element_text(size = 7),
        axis.ticks = element_line(color = "black", size = 0.25),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 6))
```

Annotate GTP and ATP binding proteins

```{r}
all_atp_binder <- AnnotationDbi::select(
  org.Hs.eg.db::org.Hs.eg.db, 
  keys = "GO:0005524", 
  columns = c("SYMBOL", "IPI"), 
  keytype = "GOALL")

all_gtp_binder <- AnnotationDbi::select(
  org.Hs.eg.db::org.Hs.eg.db, 
  keys = "GO:0005525", 
  columns = c("SYMBOL", "IPI"), 
  keytype = "GOALL")
```


Download the supplementary table from the journal's website

```{r}
# still needs to be added when supplementary table is available online
```


Read in the data and reformat to a data frame as would be obtained after import of the raw data:

```{r}
gtp_raw <- read_xlsx("Supplementary_Data_3.xlsx", sheet = "GTP") %>% 
  dplyr::select(representative,
                clustername,
                qupm,
                qusm,
                experiment,
                temperature,
                matches("sumionarea"),
                matches("rel_fc_protein"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
gtp_fil <- resolveAmbiguousProteinNames(gtp_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
gtp_df <- recomputeSignalFromRatios(gtp_fil)
```

Compute null and alternative model fits and extract parameters
```{r eval=reRun}
gtp_params_df <- getModelParamsDf(gtp_df, maxit = 500)
saveRDS(gtp_params_df, file = "../pre_run_data/gtp_params_df.rds")
```

```{r eval=!reRun, echo=FALSE}
gtp_params_df <- readRDS("../pre_run_data/gtp_params_df.rds")
```

Compute *F* statistics
```{r}
gtp_fstat_df <- computeFStatFromParams(gtp_params_df)
```

Get $B$ datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:
```{r eval=reRun}
set.seed(12, kind = "L'Ecuyer-CMRG")
gtp_null_df <- bootstrapNullAlternativeModel(
  df = gtp_df, params_df = gtp_params_df, 
  maxit = 500, B = 100,
  BPPARAM = BiocParallel::MulticoreParam(workers = 20, progressbar = TRUE),
  verbose = FALSE)
saveRDS(jq1_null_df, file = "../pre_run_data/jq1_null_df.rds")
```

```{r eval=!reRun, echo=FALSE}
gtp_null_df <- readRDS("../pre_run_data/gtp_null_df.rds")
```

Compute FDR and find hits:

```{r}
gtp_fdr_df <- getFDR(df_out = gtp_fstat_df,
                     df_null = gtp_null_df)
  
gtp_hits_df <- findHits(gtp_fdr_df, alpha = 0.1)
```

```{r}
ggplot(gtp_fdr_df %>% 
         filter(dataset == "true"), 
       aes(log2(rssH0 - rssH1), asinh(F_statistic))) +
  geom_point(color = "gray", alpha = 0.5, size = 0.5) + 
  geom_point(color = "black", alpha = 0.5, 
             size = 0.5,
             data = gtp_hits_df %>% 
              filter(!clustername %in% 
                       all_atp_binder$SYMBOL,
                     !clustername %in% 
                       all_gtp_binder$SYMBOL)) + 
  geom_point(color = "darkgreen", alpha = 0.5, 
             size = 0.5,
             data = filter(gtp_hits_df, clustername %in% 
                             all_atp_binder$SYMBOL)) + 
  geom_point(color = "violet", alpha = 0.5, 
             size = 0.5,
             data = filter(gtp_hits_df, clustername %in% 
                             all_gtp_binder$SYMBOL)) +
  facet_wrap(~nObsRound, scales = "free", ncol = 5) +
   labs(x = expression('log'[2]~'(RSS'^0~' - RSS'^1~')'),
       y = expression('asinh('*italic(F)*' statistic)')) +
  coord_cartesian(xlim = c(-12.5, 7.5), ylim = c(0, 6)) +
  ggtitle("GTP gelfil. lysate experiment") +
  theme_paper
```

# Compare sets

```{r}
gtp_threshold_df <- read_xlsx("Supplementary_Data_3.xlsx", sheet = "GTP") %>% 
  dplyr::select(clustername, stabilized_hits_found_by_threshold) %>% 
  filter(stabilized_hits_found_by_threshold)

gtp_hits_10per <- findHits(gtp_fdr_df, 0.1) %>% 
  filter(detected_effectH1 == "stability", slopeH1 > 0)
gtp_hits_1per <- findHits(gtp_fdr_df, 0.01) %>% 
  filter(detected_effectH1 == "stability", slopeH1 > 0)


set_intersect_df <- data.frame(
  gene_name = filter(gtp_fdr_df, !duplicated(clustername))$clustername) %>% 
  mutate(`ATP binders` = as.numeric(gene_name %in% all_atp_binder$SYMBOL),
         `GTP binders` = as.numeric(gene_name %in% all_gtp_binder$SYMBOL),
         `DLPTP hits at 10% FDR` = 
           as.numeric(gene_name %in% gtp_hits_10per$clustername),
         `DLPTP hits at 1% FDR` = 
           as.numeric(gene_name %in% gtp_hits_1per$clustername),
         `Hits found by threshold-based approach` = 
           as.numeric(gene_name %in% gtp_threshold_df$clustername))

upset(set_intersect_df, nsets = 5)
```


```{r}
sessionInfo()
```

