---
title: "Benchmark on synthetic dataset"
author:
- name: Nils Kurzawa
  affiliation: 
  - European Molecular Biology Laboratory (EMBL), Genome Biology Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
package: TPP2D
output:
  BiocStyle::pdf_document:
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8} 
bibliography: ../bibliography.bib
csl: ../cell.csl
header-includes: 
- \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reRun <- FALSE
```

# Step-by-step walk through the `TPP2D` analysis

```{r eval = FALSE}
# This script uses the development version of TPP2D
if(require("BiocManager"))
  install.packages("BiocManager")
BiocManager::install("nkurzaw/TPP2D")
```

Load required libraries

```{r}
library(TPP2D)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(TPP)
```

Download and read in required datasets for true positive spike-in profiles:
- Panobinostat cell dataset @Becher2016
- JQ1 lysate dataset @Savitski2018
- Ampicillin *E. coli* lysate @Mateus2018
- ATP crude lysate rep1 @Sridharan2019


```{r}
# Pano datasets
if(!file.exists("../Panobinostat_HepG2_cell_2DTPP_Becher_et_al_2016/41589_2016_BFnchembio2185_MOESM254_ESM.xlsx")){
download.file(
  url = "https://static-content.springer.com/esm/art%3A10.1038%2Fnchembio.2185/MediaObjects/41589_2016_BFnchembio2185_MOESM254_ESM.xlsx",
  destfile = "../Panobinostat_HepG2_cell_2DTPP_Becher_et_al_2016/41589_2016_BFnchembio2185_MOESM254_ESM.xlsx") 
}

# JQ1 dataset
if(!file.exists("../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3/Supplementary Dataset 2_2D-TPP.xlsx")){
    download.file(
      url = "https://data.mendeley.com/datasets/8pzhg2tdyb/1/files/115f60c9-01d1-4213-9abb-aa095d70a626/Savitski%20et%20al%20Figure%203.zip?dl=1",
      destfile = "../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018Savitski_et_al_Figure_3.zip") 
    unzip(zipfile = "../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3.zip", 
        exdir = "../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3")
    system("rm ../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3.zip")
}

# Ampicillin E.coli dataset


# ATP crude lysate dataset
```

```{r}
jq1_lys_raw <- read_xlsx("../JQ1_THP1_lysate_2DTPP_Savitski_et_al_2018/Savitski_et_al_Figure_3/Supplementary Dataset 2_2D-TPP.xlsx", sheet = 3, skip = 1) %>% 
  dplyr::select(representative = `Accession No.`,
                clustername = `protein name`,
                qupm = QUPM,
                qusm = QUSM,
                temperature,
                matches("sumionarea"),
                -matches("total"),
                matches("rel_fc_protein"),
                -matches("transformed"),
                -matches("orig"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
jq1_lys_fil <- resolveAmbiguousProteinNames(jq1_lys_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
jq1_lys_df <- recomputeSignalFromRatios(jq1_lys_fil)
```

```{r}
pano_raw <- read_xlsx("41589_2016_BFnchembio2185_MOESM254_ESM.xlsx", 
                           sheet = 1, skip = 1) %>% 
  dplyr::select(representative,
                clustername,
                experiment = ms_experiment,
                qupm,
                qusm,
                temperature,
                matches("sumionarea"),
                -matches("total"),
                matches("rel_fc_protein"),
                -matches("transformed"),
                -matches("orig"),
                -matches("log2rel"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
pano_fil <- resolveAmbiguousProteinNames(pano_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
pano_df <- recomputeSignalFromRatios(pano_fil)
```

```{r}
atp_df <- readRDS("../pre_run_data/atp_df.rds")
amp_df <- readRDS("../pre_run_data/amp_df.rds")
```

```{r}
atp_fil <- TPP2D:::.independentFilter(atp_df, fcThres = 1.25)
```


```{r}
all_atp_binder <- AnnotationDbi::select(
  org.Hs.eg.db, 
  keys = "GO:0005524", 
  columns = c("SYMBOL", "IPI"), 
  keytype = "GOALL")
```

# Identify noise induced variation from JQ1 dataset

## JQ1 lysate
```{r}
sd_log2_val_jq1 <- jq1_lys_df %>% 
  group_by(representative, temperature) %>% 
  summarise(sd_log2_value = sd(log2_value)) %>% 
  ungroup() %>% 
  filter(!is.na(sd_log2_value)) %>% 
  group_by(temperature) %>% 
  mutate(quant_95 = quantile(sd_log2_value, 0.95)) %>% 
  filter(sd_log2_value < quant_95) %>% 
  ungroup()

ggplot(sd_log2_val_jq1, aes(sd_log2_value)) +
  geom_density()+
  facet_wrap(~temperature)
```


# Simulate data

```{r eval=reRun}
set.seed(12)
nobs <- c(20, 30, 40, 50, 60)

log_concs <- c(-Inf, -7, -6, -5.3, -4.7)

temperatures <- c(42.1, 44.1, 46.2, 48.1, 
                  50.4, 51.9, 54.0, 56.1,
                  58.2, 60.1, 62.4, 63.9)

alphas <- seq(0, 1, 0.01)

beta_max <- seq(2, 6, by = 0.01)

ec_50s <- seq(-7, -5, by = 0.01)

slopes <- c(rnorm(500, mean = -1, sd = 0.5),
  rnorm(500, mean = 1, sd = 0.5))


prots <- paste("protein", 1:5000, sep = "")

log2_value_baselines <- seq(21, 27, by = 0.1)

sim_df <- bind_rows(lapply(prots, function(pr){
  # draw number of observations and define temperatures 
  nobs_p <- sample(nobs, 1)
  temp_p <- temperatures[seq_len(nobs_p/5)]
  
  # sample base level for each temperature
  base_p <- sort(sapply(temp_p, function(x) sample(log2_value_baselines, 1)), 
                 decreasing = TRUE)
  # add noise for each concentration
  log2_values <- unlist(lapply(seq_len(length(temp_p)), function(i){
    sd_log2_inten <- 5
    while(sd_log2_inten > max(filter(
      sd_log2_val_jq1, temperature == temp_p[i])$sd_log2_value)){
        log2_inten <- 
          rnorm(5, mean = base_p[i], 
                sd = base::sample(filter(
                  sd_log2_val_jq1, temperature == temp_p[i])$sd_log2_value, 1))
        sd_log2_inten <- sd(log2_inten)
    }
    return(log2_inten)

    }))
  # assemble df
  prot_df <- tibble(clustername = pr,
                    temperature = rep(temp_p, each = 5),
                    log_conc = rep(log_concs, length(temp_p)),
                    log2_value = log2_values,
                    raw_value = 2^log2_values)
})) %>% group_by(clustername, temperature) %>% 
  mutate(rel_value = raw_value / raw_value[which(log_conc == -Inf)]) %>% 
  ungroup() 


ms_exp_df <- tibble(
  ms_experiment = as.character(
    rep(seq_len(length(temperatures)/2), each = 2)),
  temperature = temperatures)   

sim_df <- left_join(sim_df, ms_exp_df) %>% 
  mutate(representative = clustername)

# spike in TPs from different datasets
sim_tp_df <- bind_rows(
  sim_df %>% dplyr::select(representative, clustername, temperature, log_conc, log2_value,rel_value),
  filter(pano_df, clustername %in% 
             c("HDAC2", "HDAC1", "PAH", "FADS1", "TTC38", 
               "FADS2", "TBP", "SCCPDH", "HDAC6")) %>% 
    filter(representative != "IPI00647298.2") %>%  
    dplyr::select(representative, clustername, temperature, log_conc, log2_value, rel_value),
  filter(jq1_df, clustername %in% 
             c("BRD2", "BRD3", "BRD4")) %>% 
    dplyr::select(representative, clustername, temperature, log_conc, log2_value, rel_value),
  filter(amp_df, clustername %in% 
             c("FTSI", "MRCA", "AMPC", "DACB", "MRCB")) %>% 
    dplyr::select(representative, clustername, temperature, log_conc, log2_value, rel_value),
  filter(atp_fil, clustername %in% 
             filter(all_atp_binder, EVIDENCEALL == "IDA")$SYMBOL) %>% 
    dplyr::select(representative, clustername, temperature, log_conc, log2_value, rel_value)
 )

# filter out proteins that will be dropped by the downstream analysis because of too few qupms or too few observations.
sim_tp_df <- sim_tp_df %>%
  filter(!clustername %in% c("IGHMBP2", "LATS1", "MAP4K2", "MTPAP", "RECQL4",
                             "SRPK1", "SRPK2", "AKT1", "DHX9", "INSR"))

saveRDS(sim_tp_df, file = "../pre_run_data/sim_tp_df.rds")
```

```{r eval=!reRun}
sim_tp_df <- readRDS("../pre_run_data/sim_tp_df.rds")
```

Compute null and alternative model fits and extract parameters
```{r eval=reRun}
sim_params_df <- getModelParamsDf(sim_tp_df, maxit = 500)
saveRDS(sim_params_df, file = "../pre_run_data/sim_params_df.rds")
```

```{r eval=!reRun, echo=FALSE}
sim_params_df <- readRDS("../pre_run_data/sim_params_df.rds")
```

Compute *F* statistics
```{r}
sim_fstat_df <- computeFStatFromParams(sim_params_df)
```

Do 3 separate rounds of bootstrapping the null distribution:
```{r eval=reRun}
set.seed(1, kind = "L'Ecuyer-CMRG")
sim_s1_null_df <- bootstrapNullAlternativeModel(
  df = sim_tp_df, params_df = sim_params_df, 
  maxit = 500, B = 100,
  BPPARAM = BiocParallel::MulticoreParam(workers = 20, progressbar = TRUE),
  verbose = FALSE)
saveRDS(sim_s1_null_df, file = "../pre_run_data/sim_s1_null_df.rds")
```

```{r eval=reRun}
set.seed(2, kind = "L'Ecuyer-CMRG")
sim_s2_null_df <- bootstrapNullAlternativeModel(
  df = sim_tp_df, params_df = sim_params_df, 
  maxit = 500, B = 100,
  BPPARAM = BiocParallel::MulticoreParam(workers = 20, progressbar = TRUE),
  verbose = FALSE)
saveRDS(sim_s2_null_df, file = "../pre_run_data/sim_s2_null_df.rds")
```

```{r eval=reRun}
set.seed(3, kind = "L'Ecuyer-CMRG")
sim_s3_null_df <- bootstrapNullAlternativeModel(
  df = sim_tp_df, params_df = sim_params_df, 
  maxit = 500, B = 100,
  BPPARAM = BiocParallel::MulticoreParam(workers = 20, progressbar = TRUE),
  verbose = FALSE)
saveRDS(sim_s3_null_df, file = "../pre_run_data/sim_s3_null_df.rds")
```

```{r eval=!reRun}
sim_s1_null_df <- readRDS("../pre_run_data/sim_s1_null_df.rds")
sim_s2_null_df <- readRDS("../pre_run_data/sim_s2_null_df.rds")
sim_s3_null_df <- readRDS("../pre_run_data/sim_s3_null_df.rds")
```

Define function to compute the precision and recall based on the simulated dataset:

```{r}
computePrecisionRecall <- function(fdr_df,
                                   alpha_range = seq(0.01, 1, by = 0.01),
                                   ntp = 50){
  out_pr_df <- bind_rows(tibble(precision = 1, recall = 0, alpha = 0),
                         bind_rows(lapply(alpha_range, function(alpha_i){
    hits_df <- TPP2D::findHits(fdr_df, alpha = alpha_i)
    
    r_df <- tibble(
        precision = length(which(!grepl("protein", hits_df$clustername)))/
            nrow(hits_df),
        recall = length(which(!grepl("protein", hits_df$clustername)))/
            ntp,alpha = alpha_i)
  })))
  
  return(out_pr_df)
}

```

Apply threshold-based approach to synthetic dataset

```{r}
# set dummy drug concentrations
general_conc_df <- tibble(
  conc_id = 1:5,
  new_conc = paste0("rel_fc_", as.character(c(0, 0.02, 0.143, 1, 5)))
)

# get example data to use temperature and experiment info
data(panobinostat_2DTPP_smallExample, package = "TPP")
config_tab <-  panobinostat_2DTPP_config
config_tab$Experiment <- rep(unique(sim_tp_df$experiment), each = 2)

general_temperature_df <- tibble(
  temp_id = 1:12,
  temperature = config_tab$Temperature
)

# generate data frame as obtained by TPP package
thres_approach_df <- sim_tp_df %>% 
  group_by(clustername, temperature) %>% 
  mutate(conc_id = dense_rank(log_conc)) %>% 
  ungroup %>% 
  left_join(general_conc_df, by = "conc_id") %>% 
  dplyr::select(-log_conc, -log2_value, -conc_id) %>% 
  group_by(clustername) %>% 
  mutate(temp_id = dense_rank(temperature)) %>% 
  dplyr::select(-temperature) %>% 
  left_join(general_temperature_df, by = "temp_id") %>% 
  spread(new_conc, rel_value) %>% 
  rowwise() %>% 
  mutate(unique_ID = paste(c(experiment, temperature, representative), collapse = "_"),
         qupm = 5) %>% 
  ungroup

importSettings <- attr(thres_approach_df, "importSettings")
importSettings$fcStr <- "rel_fc_"
importSettings$proteinIdCol <- "representative"
importSettings$uniqueIdCol <- "unique_ID"
importSettings$nonZeroCols <- "qupm"
importSettings$configTable <- as.data.frame(config_tab)
attr(thres_approach_df, "importSettings") <- importSettings

# normalize fold changes
thres_approach_norm <- tpp2dNormalize(data = thres_approach_df)

# adapt curve fit function to work with synthetic data
tpp2dCurveFitAdapted <- function (configFile = NULL, data, 
                                  nCores = 1, naStrs = NULL, 
    fcStr = NULL, idVar = NULL, nonZeroCols = NULL, r2Cutoff = 0.8, 
    fcCutoff = 1.5, slopeBounds = c(1, 50), fcTolerance = 0.1) 
{
    if (!missing(configFile)) {
        warning("`configFile` is deprecated.", call. = TRUE)
    }
    if (!missing(naStrs)) {
        warning("`naStrs` is deprecated.", call. = TRUE)
    }
    if (!missing(fcStr)) {
        warning("`fcStr` is deprecated.", call. = TRUE)
    }
    if (!missing(idVar)) {
        warning("`idVar` is deprecated.", call. = TRUE)
    }
    if (!missing(nonZeroCols)) {
        warning("`nonZeroCols` is deprecated.", call. = TRUE)
    }
    TPP:::checkFunctionArgs(match.call(), c("data"))
    importSettings <- attr(data, "importSettings")
    configTable <- importSettings$configTable
    uniqueIdCol <- importSettings$uniqueIdCol
    if (is.null(uniqueIdCol)) {
        stop("attr(data, 'uniqueIdCol') must contain a field named 'uniqueIdCol'.")
    }
    else if (!is.character(uniqueIdCol)) {
        stop("attr(data, 'importSettings')$uniqueIdCol must be of class character.")
    }
    else {
        message("Looking for unique ID column: '", uniqueIdCol, 
            "'")
    }
    if (!uniqueIdCol %in% colnames(data)) {
        stop("Please specify an uniqueIdCol character string argument that represents a suffix of one of \n         the column names of your data!")
    }
    else if (length(data[[uniqueIdCol]]) != length(unique(data[[uniqueIdCol]]))) {
        stop("Please indicate an uniqueIdCol character string that matches a column with unique identifiers!")
    }
    nonZeroCols <- importSettings$nonZeroCols
    if (is.null(nonZeroCols)) {
        stop("attr(data, 'importSettings') must contain a field named 'nonZeroCols'.")
    }
    else if (!is.character(nonZeroCols)) {
        stop("attr(data, 'importSettings')$nonZeroCols must be of class character.")
    }
    else {
        message("Looking for nonZeroCols: '", nonZeroCols, "'")
    }
    if (!all(nonZeroCols %in% colnames(data))) {
        stop("The given QC columns (specified by attr(data, 'importSettings')$nonZeroCols) were not found in the column names of 'data'.")
    }
    finalFcPrefix <- TPP:::obtain_fcStr_from_df_annotation(dat = data)
    message("Performing TPP-CCR dose response curve fitting and generating result table...")
    cfgIn <- TPP:::convert_2D_cfgTable_to_CCR_cfgTable(configTable = configTable)
    datIn <- as.data.frame(data)
    CCRresult <- suppressMessages(analyzeTPPCCR(configTable = cfgIn, 
        data = datIn, nCores = nCores, resultPath = NULL, plotCurves = FALSE, 
        fcStr = finalFcPrefix, naStrs = c("NA", "n/d", "NaN", 
            "<NA>"), qualColName = "qupm", xlsxExport = FALSE, 
        idVar = uniqueIdCol, nonZeroCols = nonZeroCols, normalize = FALSE, 
        r2Cutoff = r2Cutoff, fcCutoff = fcCutoff, slopeBounds = slopeBounds, 
        fcTolerance = fcTolerance, ggplotTheme = NULL, verbose = FALSE))
    compound <- as.character(cfgIn$Experiment)
    allCols <- colnames(CCRresult)
    newCols <- sub(paste("*_", compound, sep = ""), "", allCols)
    colnames(CCRresult) <- newCols
    message("Done.")
    importSettings$r2Cutoff <- r2Cutoff
    importSettings$fcCutoff <- fcCutoff
    importSettings$slopeBounds <- slopeBounds
    importSettings$fcTolerance <- fcTolerance
    importSettings$uniqueIdCol <- "Protein_ID"
    attr(CCRresult, "importSettings") <- importSettings
    attr(CCRresult, "configTable") <- configTable
    return(CCRresult)
}

ccr2dResults <- tpp2dCurveFitAdapted(data = thres_approach_norm)
importSettings <- attr(ccr2dResults, "importSettings")
importSettings$intensityStr <- "rel_fc_"
attr(ccr2dResults, "importSettings") <- importSettings

passed_filter_df <- ccr2dResults %>% tbl_df %>% filter(passed_filter)
thres_found_recurrently <- unique(passed_filter_df$clustername[
    duplicated(passed_filter_df$clustername)])
```



Get FDR estimates for different numbers of bootstraps
```{r}
sim_fdr_df_s1_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_null_df)
sim_fdr_df_s2_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s2_null_df)
sim_fdr_df_s3_b100_byMsExp <- getFDR(df_out = sim_fstat_df,
                                     df_null = sim_s3_null_df)

sim_fdr_df_s1_b50_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:50, sep = "_")))
sim_fdr_df_s1_b50_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:50, sep = "_")))
sim_fdr_df_s1_b50_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:50, sep = "_")))

sim_fdr_df_s1_b20_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:20, sep = "_")))
sim_fdr_df_s2_b20_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s2_null_df, dataset %in% paste("bootstrap", 1:20, sep = "_")))
sim_fdr_df_s3_b20_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s3_null_df, dataset %in% paste("bootstrap", 1:20, sep = "_")))

sim_fdr_df_s1_b10_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:10, sep = "_")))
sim_fdr_df_s2_b10_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s2_null_df, dataset %in% paste("bootstrap", 1:10, sep = "_")))
sim_fdr_df_s3_b10_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s3_null_df, dataset %in% paste("bootstrap", 1:10, sep = "_")))

sim_fdr_df_s1_b5_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_null_df, dataset %in% paste("bootstrap", 1:5, sep = "_")))
sim_fdr_df_s2_b5_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s2_null_df, dataset %in% paste("bootstrap", 1:5, sep = "_")))
sim_fdr_df_s3_b5_byMsExp <- getFDR(df_out = sim_fstat_df,
                         df_null = filter(sim_s3_null_df, dataset %in% paste("bootstrap", 1:5, sep = "_")))
```

Compute precision and recall for each of the FDR estimates based on the different number of bootstraps
```{r}
pr_b100_1_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s1_b100_byMsExp, ntp = 80)
pr_b100_2_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s2_b100_byMsExp, ntp = 80)
pr_b100_3_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s3_b100_byMsExp, ntp = 80)

pr_b20_1_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s1_b20_byMsExp, ntp = 80)
pr_b20_2_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s2_b20_byMsExp, ntp = 80)
pr_b20_3_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s3_b20_byMsExp, ntp = 80)

pr_b5_1_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s1_b5_byMsExp, ntp = 80)
pr_b5_2_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s2_b5_byMsExp, ntp = 80)
pr_b5_3_byMsExp <- computePrecisionRecall(
    fdr_df = sim_fdr_df_s3_b5_byMsExp, ntp = 80)

all_pr_df <- bind_rows(pr_b5_1_byMsExp %>% 
                         mutate(B = "5"), 
                       pr_b5_2_byMsExp %>% 
                         mutate(B = "5"), 
                       pr_b5_3_byMsExp %>% 
                         mutate(B = "5"), 
                       pr_b20_1_byMsExp %>% 
                         mutate(B = "20"),
                       pr_b20_2_byMsExp %>% 
                         mutate(B = "20"),
                       pr_b20_3_byMsExp %>% 
                         mutate(B = "20"),
                       pr_b100_1_byMsExp %>% 
                         mutate(B = "100"),
                       pr_b100_2_byMsExp %>% 
                         mutate(B = "100"),
                       pr_b100_3_byMsExp %>% 
                         mutate(B = "100"),
                       tibble(
                         precision = length(thres_found_recurrently[!grepl("protein", thres_found_recurrently)])/
                             length(thres_found_recurrently),
                         recall = length(thres_found_recurrently[!grepl("protein", thres_found_recurrently)])/80,
                         alpha = 0,
                         B = "Threshold-based approach")) %>% 
  group_by(B, alpha) %>% 
  summarize(mean_fdr = mean(1 - precision),
            mean_tpr = mean(recall),
            sd_fdr = sd(1 - precision),
            sd_tpr = sd(recall)) %>% 
  ungroup %>% 
  mutate(zoom_in = case_when(
      mean_fdr < 0.15 & mean_tpr > 0.55 & mean_tpr < 0.85 ~ TRUE,
      TRUE ~ FALSE))


roc_main <- ggplot(all_pr_df, aes(x = mean_fdr, y = mean_tpr, color = as.factor(B))) + 
  geom_path(alpha = 0.65, 
            size = 0.25) + 
  geom_point(shape = 8,
             data = filter(filter(all_pr_df, alpha == "0.1"))) +
    geom_point(shape = 3, 
             data = filter(filter(all_pr_df, alpha == "0.05"))) +
    geom_point(shape = 4,
             data = filter(filter(all_pr_df, alpha == "0.01"))) +
  geom_point(shape = 6,
             data = filter(filter(all_pr_df, B == "Becher et al."))) +
  scale_color_brewer("B", palette = "Set1") +
  labs(x = "FDR", y = "TPR") +
  geom_vline(xintercept = c(0.01, 0.05, 0.1),
             linetype = "dashed", alpha = 0.65,
             size = 0.25) +
  geom_rect(xmin = -0.01, xmax = 0.15, 
            ymin = 0.55, ymax = 0.85,
            #alpha = 0.5, 
            color = "grey90",
            data = tibble(mean_fdr = 1,
                          mean_tpr = 1,
                          B = 1),
            fill = NA) +
  coord_fixed() +
  theme_classic() 


roc_zoom <- ggplot(all_pr_df, 
       aes(x = mean_fdr, y = mean_tpr, color = as.factor(B))) + 
  geom_path(alpha = 0.65, 
            size = 0.25) + 
  geom_point(shape = 8,
             data = filter(filter(all_pr_df, alpha == "0.1"))) +
    geom_point(shape = 3, 
             data = filter(filter(all_pr_df, alpha == "0.05"))) +
    geom_point(shape = 4,
             data = filter(filter(all_pr_df, alpha == "0.01"))) +
  geom_point(shape = 6,
             data = filter(filter(all_pr_df, B == "Becher et al."))) +
  scale_color_brewer("B", palette = "Set1") +
  labs(x = "FDR", y = "TPR") +
  geom_vline(xintercept = c(0.01, 0.05, 0.1),
             linetype = "dashed", alpha = 0.65,
             size = 0.25) +
  coord_fixed(xlim = c(0, 0.15), ylim = c(0.55, 0.85)) +
  theme_classic() +
  theme(legend.position = "none")


cowplot::plot_grid(roc_zoom, roc_main, rel_widths = c(0.5,1))
```


```{r}
sessionInfo()
```

