---
title: "Analysis of BRD-3811 dataset"
author:
- name: Nils Kurzawa
  affiliation: 
  - European Molecular Biology Laboratory (EMBL), Genome Biology Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
package: TPP2D
output:
  BiocStyle::pdf_document:
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8} 
bibliography: ../bibliography.bib
csl: ../cell.csl
header-includes: 
- \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reRun <- FALSE
```

# Step-by-step walk through the `TPP2D` analysis

```{r eval = FALSE}
# This script uses the development version of TPP2D
if(require("BiocManager"))
  install.packages("BiocManager")
BiocManager::install("nkurzaw/TPP2D")
```

Load required libraries

```{r}
library(TPP2D)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(clusterProfiler)
library(org.Hs.eg.db)
```

Define plot style
```{r}
theme_paper <- theme_bw(base_size = 6) +
  theme(legend.background = element_blank(), 
        legend.key = element_blank(), 
        panel.background = element_blank(), 
        panel.grid.major = element_line(colour = "grey92", size = 0.25),
        panel.grid.minor = element_line(colour = "grey92", size = 0.15),
        panel.border = element_blank(), 
        strip.background = element_blank(), 
        plot.background = element_blank(), 
        complete = TRUE,
        axis.line = element_line(color = "black", size = 0.25),
        text = element_text(size = 7),
        axis.ticks = element_line(color = "black", size = 0.25),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 6))
```

Download the supplementary table from the journal's website

```{r}
# still needs to be added when supplementary table is available online
```


Read in the data and reformat to a data frame as would be obtained after import of the raw data:

```{r}
brd_raw <- read_xlsx("Supplementary_Data_2.xlsx", sheet = "BRD3811") %>% 
  dplyr::select(representative,
                clustername,
                qupm,
                qusm,
                temperature,
                matches("sumionarea"),
                matches("rel_fc_protein"))  %>% 
  gather(key, value, matches("sumionarea"), matches("rel_fc_protein")) %>% 
  mutate(conc = as.numeric(gsub("uM", "", gsub(".+_protein_[0-9,H,L]+_[0-9,H,L]+_", "", key))),
         temperature = as.numeric(gsub("C", "", temperature)),
         key = case_when(grepl("sumionarea", key) ~ "raw_value",
                         grepl("rel_fc", key) ~ "rel_value")) %>% 
  spread(key, value) %>% 
  arrange(representative, temperature, conc) %>% 
  group_by(clustername, temperature, conc) %>% 
  filter(qupm == max(qupm), 
         qusm == max(qusm), 
         raw_value == max(raw_value)) %>% 
  filter(!duplicated(clustername)) %>% 
  ungroup %>% 
  mutate(log2_value = log2(raw_value),
         log_conc = log10(conc/1e6)) %>% 
  filter(qupm > 1)

# resolve ambiguous protein names
brd_fil <- resolveAmbiguousProteinNames(brd_raw)
  
# recompute reporter ion signal from robust Isobarquant fold changes
brd_df <- recomputeSignalFromRatios(brd_fil)
```

Compute null and alternative model fits and extract parameters
```{r eval=reRun}
brd_params_df <- getModelParamsDf(brd_df, maxit = 500)
saveRDS(brd_params_df, file = "../pre_run_data/brd_params_df.rds")
```

```{r eval=!reRun, echo=FALSE}
brd_params_df <- readRDS("../pre_run_data/brd_params_df.rds")
```

Compute *F* statistics
```{r}
brd_fstat_df <- computeFStatFromParams(brd_params_df)
```

Get $B$ datasets expected under the null model and perform model fitting and compute F statistics to obtain a null distribution for FDR calibration:
```{r eval=reRun}
set.seed(12, kind = "L'Ecuyer-CMRG")
brd_null_df <- bootstrapNullAlternativeModel(
  df = brd_df, params_df = brd_params_df, 
  maxit = 500, B = 100,
  BPPARAM = BiocParallel::MulticoreParam(workers = 20, progressbar = TRUE),
  verbose = FALSE)
saveRDS(brd_null_df, file = "../pre_run_data/brd_null_df.rds")
```

```{r eval=!reRun, echo=FALSE}
brd_null_df <- readRDS("../pre_run_data/brd_null_df.rds")
```

Remove carry-over cases:
```{r}
## manually identified carry-over cases
carry_over_cases <- c("NQO1", "CSK")

```


Compute FDR and find hits:

```{r}
brd_fdr_df <- getFDR(df_out = brd_fstat_df %>% 
                         filter(!clustername %in% carry_over_cases),
                     df_null = brd_null_df %>% 
                         filter(!clustername %in% carry_over_cases))
  
brd_hits_df <- findHits(brd_fdr_df, alpha = 0.1)
```

```{r}
ggplot(brd_fdr_df %>% 
         filter(dataset == "true"), 
       aes(log2(rssH0 - rssH1), asinh(F_statistic))) +
  geom_point(color = "gray", alpha = 0.5, size = 1) + 
  geom_point(aes(color = group), alpha = 0.5, 
             size = 1,
             data = brd_hits_df %>% 
                 mutate(group = case_when(
                     slopeH1 > 0 ~ "stabilized protein",
                     slopeH1 < 0 ~ "destabilized protein"))) + 
  ggrepel::geom_text_repel(
    aes(label = clustername),
    data = filter(brd_hits_df, clustername %in% 
                    c("HDAC8", "LAP3")),
    size = 2, segment.size = 0.2, min.segment.length = unit(1, "pt")) +
  scale_color_manual("", values = c("orange", "steelblue")) +
  coord_cartesian(xlim = c(-12.5, 7.5)) + 
  ylab("asinh(F statistic)") +
  labs(x = expression('log'[2]~'(RSS'^0~' - RSS'^1~')'),
       y = expression('asinh('*italic(F)*' statistic)')) +
  ggtitle("brd-34051 in-cell experiment") +
  theme_paper +
  theme(legend.position = "bottom")

```

# Plot example profiles

LAP3
```{r}
lap3_fit <- plot2dTppFit(brd_df, "LAP3", "H1")$data

lap3_df <- filter(brd_df, clustername == "LAP3")

ggplot(lap3_fit, aes(log_conc, y_hat)) +
  geom_line() +
  geom_point(aes(log_conc, log2_value), 
             data = lap3_df, size = 0.5) +
  facet_wrap(~temperature, ncol = 6) +
  labs(x = expression('log'[10]~ '(brd-34051 conc.)'),
       y = expression('log'[2]~ '(summed ion area)')) +
  ggtitle("LAP3, alternative model") +
  theme_paper
```

# GO analysis
We do this using the `clusterProfiler` package by @Yu2012
```{r}
hits_entrez <- bitr(brd_hits_df$clustername, 
                fromType = "SYMBOL",
                toType = c("ENTREZID"),
                OrgDb = org.Hs.eg.db)

backg_entrez <- bitr(brd_fdr_df$clustername, 
                     fromType = "SYMBOL",
                     toType = c("ENTREZID"),
                     OrgDb = org.Hs.eg.db)


ego <- enrichGO(gene = hits_entrez$ENTREZID,
                universe = backg_entrez$ENTREZID,
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05,
                readable = TRUE)

dotplot(ego)
```


```{r}
sessionInfo()
```
